Symmetric Circuits for Rank Gates

### 2018-04-06 sections/conclusion.tex
Most recent changes to the conclusion, with edits all around the paper.

The main result of this paper establishes an equivalence between the expressive
power of $\FPR$ and the computational power of $\PT$-uniform transparent
symmetric rank-circuits, generalising a similar result by Anderson and
Dawar~\cite{AndersonD17}. Our motivation for proving this characterisation  The hope is that this characterisation will
contribute to our understanding of one of the two candidate logics for capturing
$\PT$.

% In the quest for a logic that captures polynomial time

% At the time of writing there are two candidate logics which Recent work on the
% problem of capturing $\PT$ has produced two `candidate' logics for capturing
% polynomial-time. In this paper we extend Anderson and Dawar's symmetric
% circuit characterisation of $\FPC$, showing that $P$-uniform families of
% transparent symmetric circuits with rank gates have the same expressive power
% as $\FPR$. Since $\FPR$ remains one of the two c the candidate logics for
% capturing $\PT$, along with choiceless polynomial time,
\end{document}

>>>>

$\FPR$ is one of the two candidate logics for capturing $\PT$ an so
understanding its expressive power is an absolutely central problem in the
field. The main result of this paper establishes an equivalence between the
expressive power of $\FPR$ and the computational power of $\PT$-uniform families
of transparent symmetric rank-circuits. Not only does this result establish an
interesting characterisation of a little-understood and crucially important
logic, but it also further develops our understanding of the connection between
logic and circuit complexity, allowing our understanding of $\FPR$ (and
circuits) to benefit from ideas in both of these quite disparate fields.

This alternative view of the logic also helps emphasise certain characteristics
of the logic that may be much less obvious when working with the conventional
definition. To choose one notable example, the circuit characterisations (both
of $\FPR$ and $\FPC$) help bring to the fore the importance of the symmetry
properties inherent in the the syntactic structure of the logic. Indeed, we
recall that $\PT$-uniform families of transparent invariant (rather than
symmetric) rank-circuits and $\PT$-uniform families of invariant circuits with
symmetric gates decide exactly those queries that are in $\PT$. In this way the
restriction from families of invariant circuits to families of symmetric
circuits corresponds to the restriction from $\PT$ to $\FPR$ or $\FPC$. In this
sense, we might say that it is this underlying symmetry property of the logic,
brought into focus by the circuit characterisation, that separates $\FPC$ from
$\PT$, and which determines the relationship between $\FPR$ and $\PT$. In fact,
we have that $\FPR$ captures $\PT$ if, and only if, $\PT$-uniform families of
symmetric rank-circuits have the same computational power as $\PT$-uniform
families of invariant rank-circuits. In this way the circuit-characterisation
brings out this underlying symmetry property of the logic, demonstrates how
crucial this symmetry property for understanding the logic, and then provides us
with a framework for studying symmetry.

% This illustrates just one important way in which this circuit characterisation
% aids in our understanding of $\FPR$, and perhaps in our understanding of
% logics more generally.

% In this way the circuit-characterisation allows us to under would seem then
% that understanding these symmetry properties of the circuit (and logic) is
% central to understanding the relationship between $\FPR$ (or perhaps any
% logic) and $\PT$. As such, we might thus rephrase the question of whether
% $\FPR$ captures $\PT$ in a slightly informal way as, `Does symmetry matter for
% rank-circuits?'.

However, much of the contribution of this paper is not in the main result
itself, but rather in the entirely novel framework for circuits and the numerous
new results that we needed to develop in order to prove the main result of this
paper. As such, we shall briefly review some of these contributions, placing an
emphasis both on the nature of these contributions and on the necessity of their
introduction. Indeed, it would seem at first glance that generalising Anderson
and Dawar's circuit characterisation of $\FPC$ to $\FPR$ should be a simple
matter of extending the model so as to allow circuits to be defined over bases
that includes rank-threshold functions -- and perhaps extending a few key
results and definitions as needed. We showed in
Theorem~\ref{thm:symmetric-circuits-bound} that, in fact, we cannot increase the
expressive power of the model by adding symmetric functions to the basis, and so
the addition of non-symmetric functions to the basis seems almost unavoidable.
However, while the inclusion of rank-threshold functions in the basis may seem
benign, it in fact violates a near-ubiquitous assumption in circuit complexity,
namely that the basis over which the circuit is defined must consist of
symmetric functions. While this assumption is usually not stated, the fact that
the circuit is taken to be a directed acyclic graph means that there is no order
(or any structure at all) on the inputs of each gate, an assumption which only
makes sense if it is assumed that the evaluation of each gate is invariant under
any ordering of its inputs -- which is to say the basis must be symmetric.
However, we found that in Anderson and Dawar's case this assumption pervades not
just their definition of a circuit, but also the central ideas of their paper,
including their definition of a circuit automorphism and of a symmetric circuit,
as well as almost all of the key results and techniques they use to establish
their characterisation.

As such, in order to include rank-threshold functions in the basis we needed to
construct a new framework for circuits as well as develop novel approaches and
techniques in order to work with these circuits. We began by generalising the
notion of a symmetric function, introducing isomorphism-invariant structured
functions. These functions take in structures encoded and strings and are
invariant under isomorphism, rather than under arbitrary permutations of the
string. It is easy to see that rank-threshold functions are
isomorphism-invariant. We defined a basis to be a set of isomorphism-invariant
functions, and extended the circuit model, adding to each gate in the circuit a
structure on its input gates in accord with the isomorphism-invariant structured
function labelling that gate. Rather than a gate being invariant under arbitrary
permutations of its input gates, a gate in this circuit should be invariant
under permutations of its input gate that form a (labelled) isomorphism. We
extended the notion of an automorphism to reflect this and defined what it means
for a circuit to be symmetric.

While we have generalised a few results for this framework, including the
support theorem, it is important to note that a number of crucial counting
techniques and polynomial-time algorithms for circuits are either not easy to
generalise or otherwise resist generalisation entirely. This follows from the
fact that a circuit automorphism needs to preserve more structure when acting on
non-symmetric gates than when acting on symmetric gates. As a result, while many
symmetry-related circuit properties are polynomial-time decidable for circuits
with symmetric gates, we have shown that these same properties are at least as
hard as the graph-isomorphism for circuits with non-symmetric gates. This is a
problem as the polynomial-time decidability of these properties is essential for
our translation from circuits into formulas.

% used throughout this paper, and is needed in order to apply the support
% theorem or reason about orbits, we need to

% for are we have shown that many natural circuit properties which are known to
% be polynomial-time decidable for circuits over symmetric bases are at least as
% hard to decide for circuits over non-symmetric bases as the graph isomorphism
% problem.

% This follows from in a rather subtle way from the fact that
% syntactic-equivalence for symmetric gates requires nothing more than checking
% if two quotiented sets are equal, while syntactic-equivalence between
% non-symmetric gates involves checking if two quotiented structures are equal.
% As such, while determining many

% While for circuits with functions into the basis. In Anderson and Dawar's case
% establishing the image of a gate under the action of an automorphism is
% entirely determined by the

% In particular, in order to define the translation from families of circuits to
% formulas we need that a number However, in order to prove this result we need
% to be able to compute in polynomial-time many basic symmetry-related
% properties of the circuit, and most of these properties reference the
% existence of an isomorphism between the structures associated with each gate.
% We show in Section~\ref{sec:transparent} that, in fact, deciding many of these
% properties for (general) rank-circuits is at least as hard as the
% graph-isomorphism problem. Importantly, the circuits studied by Anderson and
% Dawar have all of the relevant algorithmic, as in this case the isomorphism
% problems that arises reduces to set-equivalence, which is obviously
% polynomial-time decidable.

% In partoTo be more procides This is a problem, as these algorithmic properties
% are needed throughout the paper

% In particular, while almost all of the relevant properties are polynomial-time
% decidable for circuits with symmetric gates, enabling a in polynomial-time
% transformed into a normal form with unique extensions. Moreover, this normal
% form also many other crucial algorithmic properties needed to prove the main
% result, particularly the ability to compute the action of a permutation on a
% circuit, in polynomial-time. We showed in Section~\ref{} that these useful
% algorithmic properties are not (necessarily) preserved by the generalisation.
% In particular, we showed that none of these properties are polynomial-time
% decidable for rank-circuits (or any class of circuits over a non-symmetric
% basis) unless the graph isomorphism is in $\PT$.

% In order to prove our main result we $\PT$-uniform families of symmetric
% rank-circuits into equivalent formulas of $\FPR$ and visa

% Dawar. However, However, It would be natural to take symmetric rank-circuits
% as the natural generalisation of Anderson and Dawar's symmetric circuits and
% aim to prove an equivalence between the expressive power $\FPR$ and the
% computational power of $\PT$-uniform families symmetric rank-circuits.
% However, in order to prove this result we need to be able to compute in
% polynomial-time many basic symmetry-related properties of the circuit, and
% most of these properties reference the existence of an isomorphism between the
% structures associated with each gate. We show in Section~\ref{sec:transparent}
% that, in fact, deciding many of these properties for (general) rank-circuits
% is at least as hard as the graph-isomorphism problem. Importantly, the
% circuits studied by Anderson and Dawar have all of the relevant algorithmic,
% as in this case the isomorphism problems that arises reduces to
% set-equivalence, which is obviously polynomial-time decidable.

% In order to ensure

% when we introduce rank-gates this isomorphism problem that computing the
% action of an automorphism, or showing that a circuit is symmetric, or
% computing a great number of important properties is at least as hard for
% rank-circuits as the graph-isomorphism problem.

% to be able to apply the support theorem to circuits, and so needed a normal
% form for circuits with unique extensions, and we also needed to ensure that a
% number of basic circuit properties, and particularly those relating to
% symmetry, are polynomial-time decidable to circuits as well as compute a
% number of basic circuit properties in polynomial-time. This is not a problem
% for Anderson and Dawar, as all of the relevant circuit properties are
% polynomial-time decidable for symmetric-circuits with symmetric gates.
% Moreover, can be easily transformed into a normal form that has unique
% extensions (allowing for the application of the support theorem) and all of
% the nessasary circuit-properties are

% apply the support theorem, a crucial step in out proof, and in order to

% the introduction of structured functions complicates the analysis of such
% circuits. As we showed in Section~\ref{sec:transparent}, most of the
% properties of these rank-circuits are at least as hard to decide as the graph
% isomorphism problem. This poses a problem as the approach we take (and the
% approach Anderson and Dawar take) to proving this characterisation make
% repeated use of the polynomial-time decidability of these properties. Indeed,
% the two key results for translating $\PT$-uniform circuit families into
% equivalent formulas, the support theorem and the inductive evaluation of a
% gate (see Section~\ref{sec:circuits-to-formulas}), make use of polynomial-time
% algorithms for transforming a circuit into an equivalent circuit with unique
% labels, as well as for deciding if a circuit is symmetric, computing orbits,
% and deciding syntactic-equivalence.

As such, instead of working with general rank-circuits we have restricted our
attention to \emph{transparent} circuits, and proved our characterisation in
terms of $\PT$-uniform families of transparent symmetric rank-circuits. As we
have shown in Section~\ref{sec:transparent}, all of the relevant properties are
polynomial-time decidable for transparent circuits. Since all circuits with
symmetric gates are transparent, we might think of transparent circuits as an
alternative, more `algorithmically accessible', generalisation of the circuits
studied by Anderson and Dawar. However, while this restriction to transparent
circuits aids in our translation from circuits to formulas, it complicates our
translation in the other direction. Indeed, it is not hard to see that the
conventional approach for translating formulas of $\FO$ (or $\FPC$, or other
extensions of $\FO$) into $\PT$-uniform families of circuits does not, in
general, produce a family of transparent circuits. In order to address this we
defined a novel translation from formulas of $\FPR$ into equivalent
$\PT$-uniform families of transparent symmetric rank-circuits. This translation
is defined in two steps. First, we defined a translation from $\FPR$ to
$\PT$-uniform families of bounded-width formulas of $\FOrk$ and, second, we
defined a translation from these families of formulas into $\PT$-uniform
families of transparent symmetric rank-circuits. This translation, along with
the main result of the paper, gives us an equivalence between the expressive
power of three formalisms, which we summarise as follows:

% \begin{center}
%   \begin{tikzcd}
%     \FPR \ar[r, equal] \ar[dr, equal] & \text{$\PT$-uniform families of
%     bounded-width $\FOrk$-formulas}  \ar[d, equal]\\
%     & \text{$\PT$-uniform families of transparent symmetric rank-circuits}
%   \end{tikzcd}
% \end{center}

\begin{center}
  \begin{tikzcd}
    \FPR \ar[r, equal] \ar[dr, equal] & \parbox{0.35\textwidth}{$\PT$-uniform
      families of
      bounded-width $\FOrk$-formulas}  \ar[d, equal]\\
    & \parbox{0.38\textwidth}{$\PT$-uniform families of transparent symmetric
      rank-circuits}
  \end{tikzcd}
\end{center}

We already have a very similar result from from Anderson and
Dawar~\cite{AndersonD17} and Otto~\cite{Otto1996}:

\begin{center}
  \begin{tikzcd}
    \FPC \ar[r, equal] \ar[dr, equal] & \parbox{0.35\textwidth}{$\PT$-uniform
      families of
      bounded-width $\FOc$-formulas}  \ar[d, equal]\\
    & \parbox{0.35\textwidth}{$\PT$-uniform families of symmetric circuits with
      symmetric gates}
  \end{tikzcd}
\end{center}

These two sets of equivalences each give us three distinct formalisms equivalent
to $\FPR$ and $\FPC$, respectively. Importantly, each of these three formalisms
for $\FPC$ may be extended by a rank in quite distinct ways. We can extend
$\FOc$ by rank quantifiers and define $\FOrk$, we can add rank gates to
symmetric circuits with symmetric gates and define transparent symmetric
rank-circuits, and we can extend $\FPC$ by a rank operator and define $\FPR$. We
might expect these three quite different extensions to result in three
formalisms with quite different expressive power. However, the main result of
this paper along with the translation we discussed above, gives us that these
three extensions are indeed all equivalent. This provides evidence in favour of
the proposition that $\FPR$ is indeed the canonical, and most natural, extension
of $\FPC$. This is in contrast with the other rank logic in the literature, the
one introduced by Dawar et al.~\cite{Dawar09logicswith}, which was shown by
Gr\"{a}del and Pakusa~\cite{GradelP15a} to be strictly less expressive than
$\FPR$, and hence less expressive than $\PT$-uniform families of bounded-width
$\FOrk$ and $\PT$-uniform families of transparent symmetric rank-circuits.

% the cononical nature of the extension from $\FPC$ to $\FPR$. This suggests
% that the definition of $\FPR$ is indeed natural.

% 'This has quite interesting implications for the robustness of the definition
% of $\FPR$. We recall that there are really two rank logics in the literature,
% $\FPR$ and another, strictly less expressive rank logic introduced by Dawar et
% al. However, Gr\"{a}del and Pakusa~\cite{} to be strictly less expressive than
% $\FPR$

% But from these equivalences we have that these three seemingly distinct
% extensions each result in a formalism equivalent to $\FPR$. In contrast if we
% were to extend $\FPC$ by the family of rank non-uniform operators introduced
% by Dawar et al.~\cite{}, rather than the non-uniform rank operator we use
% instead, as was shown by Gr\"{a}del and Pakusa~\cite{} this logic would be
% strictly less expressive than $\FPR$, and hence less expressive than
% $\PT$-uniform families of bounded-width $\FOrk$ and $\PT$-uniform families of
% transparent symmetric rank-circuits. What is perhaps most striking is that
% extending each of these three quite different formalisms equivalent to $\FPC$
% using quite different mechanisms for rank results in a formalism equivalent to
% $\FPR$. of these formalisms by rank Since each of these extensions is defined
% quite differently, we might expect these three extensions of $\FPC$ to have
% different expressive power.

% But from the above sets of equivalences we have that all of these models have
% equivalent expressive power. In particular, $\FPR$, an extension of $\FPC$,
% has the same expressive power as other natural extensions by rank of other
% models with the same expressive power as $\FPC$. This suggests that $\FPR$ is
% the natural extension of $\FPC$ by rank operators. In contrast the rank logic
% introduced by Dawar et al.~\cite{} was shown by Gr\"{a}del and Pakusa~\cite{}
% to be strictly less expressive than $\FPR$, and hence less expressive than
% $\PT$-uniform families of bounded-width $\FOrk$ and $\PT$-uniform families of
% transparent symmetric rank-circuits. The results of this paper thus suggest
% that $\FPR$, rather than the rank logic of Dawar et al., should be considered
% the appropriate extension of $\FPC$ by rank.

% Since each of these extensions is defined quite differently, we might expect
% these three extensions of $\FPC$ to have different expressive power. But from
% the above sets of equivalences we have that all of these models have
% equivalent expressive power. In particular, $\FPR$, an extension of $\FPC$,
% has the same expressive power as other natural extensions by rank of other
% models with the same expressive power as $\FPC$. This suggests that $\FPR$ is
% the natural extension of $\FPC$ by rank operators. In contrast the rank logic
% introduced by Dawar et al.~\cite{} was shown by Gr\"{a}del and Pakusa~\cite{}
% to be strictly less expressive than $\FPR$, and hence less expressive than
% $\PT$-uniform families of bounded-width $\FOrk$ and $\PT$-uniform families of
% transparent symmetric rank-circuits. The results of this paper thus suggest
% that $\FPR$, rather than the rank logic of Dawar et al., should be considered
% the appropriate extension of $\FPC$ by rank.

\subsection{Future Work}
We have shown that $\FPR$ has the same expressive power as $\PT$-uniform
families of transparent symmetric rank-circuits. It is easy to see that
$\PT$-uniform families of symmetric rank-circuits must be at least as powerful
as $\PT$-uniform families of transparent symmetric rank-circuits. It is natural
to ask if these two formalisms actually have the same expressive power or
whether the transparency restriction does indeed restrict the expressive power
of the model. It is worth noting that if families of symmetric rank-circuits and
families of transparent symmetric-rank circuits do not have the same expressive
power then $\FPR$ does not capture $\PT$.

In this paper we have developed a framework for working with circuits over
non-symmetric bases. In fact, the first time we explicitly use make any
reference to linear algebra is in the proof of
Lemma~\ref{lem:rank-triple-equivilence}, right towards the end of the paper. It
would be natural to ask if our circuit characterisation of $\FPR$ can be
generalised in order to develop similar characterisations of other logics. It
might be good idea to start with some other extension of $\FP$ by some operator
and try to develop a similar circuit characterisation in terms $\PT$-uniform
transparent symmetric circuits over an appropriate basis. There are numerous
logics that are similar in definition to $\FPR$ but which are not known to have
the same expressive power as $\FPR$, for example $\FP$ extended by an operator
that expresses the solubility of systems of equations over rings (rather than
fields). A first step might be to develop a circuit characterisation of one of
these logics.

% If the general circuit for a strictly more powerful model then the main result
% gives us that that $\FPR$ does not capture $\PT$. We have not proven it here,
% but we believe it is possible to show that if the graph isomorphism problem is
% in $\PT$ then these classes are equivalent. If this is shown, separating these
% circuit classes would imply graph isomorphism is not in $\PT$.

% There are also natural questions regarding generalisation. In this paper we
% have characterised the expressive power of $\FPR$, an extension of $\FP$ by a
% rank operators, in terms of $\PT$-uniform families of symmetric circuits over
% the rank basis, where the rank basis is the extension of the standard basis by
% the a set of functions, each of which computed the rank operator for some
% fixed threshold and characteristic. It would be natural to ask if there is a
% much broader class of logics, perhaps some class of extensions of $\FP$, such
% that we can define for each logic in this class a natural characterisation in
% terms of $\PT$-uniform transparent symmetric circuits over some appropriate
% basis. We should note that the support theorem and most of the tools developed
% in this paper have application to circuits defined over any basis. In fact,
% the first time we explicitly use the properties of the rank function is in the
% proof of Lemma~\ref{}, right towards the end of the paper.

While we have developed a new and much more general circuit framework we have
not developed any approaches for proving lower bounds. As a first step out might
try to prove lower bounds for symmetric circuits with gates with bounded fan-in.
We might ask if it is possible to compute $\AND[3]$ using a symmetric circuit
with gates that have fan-in two.

An alternative, and perhaps more sophisticated, approach would be to try and use
pebble games, a standard tool for proving inexpressibility results in logic, to
directly prove lower bounds on circuits. Dawar~\cite{Dawar2016} has shown, via a
circuit-based argument, that the pebble games of Hella~\cite{Hella19961} can
played directly on symmetric circuits with symmetric gates without any need to
reference the logic. Dawar and Holm~\cite{DawarH2012} have defined a similar
pebble game for $\FPR$, and it would be interesting to know if an analogous
result can be established, allowing us play this pebble game directly on
symmetric rank-circuits and use it to prove lower bounds for these circuits.

We have shown that the $\PT$-uniform families of transparent symmetric
rank-circuits (resp. circuits with symmetric gates) have the same expressive
power as $\FPR$ (resp. $\FPC$), but we recall that if we remove the symmetry
restriction these circuit models express exactly those queries in $\PT$. A
natural next step, also suggested by Anderson and Dawar, would be try and weaken
the symmetry condition. One way this might be done would be to require that only
permutations from a particular subgroup of the symmetric group extend to
automorphisms of the circuit. It is easy to see that if we take this subgroup to
be the trivial subgroup then this nullifies the the symmetry requirement, and
the circuit model has the same expressive power as $\PT$, while if we take this
subgroup to be the entire symmetric group then we get back the full symmetry
requirement and hence our model has the same expressive power as the logic. As
such, this approach can be used to interpolate between the logic and $\PT$.

There is a rather easy way to develop a circuit characterisation for any logic
or complexity class with complete problems under first-order reductions. All
that is required is to extend the basis with the family of functions that decide
a complete problem and consider $\PT$-uniform families of transparent symmetric
circuits over this basis. We did not take this approach in our characterisation
of $\FPR$, preferring to use a more `natural' basis that more closely resembles
the actual definition of the logic. We should like to ask if there is a similar
circuit characterisation of $\CPTC$. One might start with bounded-rank fragments
of $\CPTC$ and see if these programs can be translated into circuits. We should
note that it does seem as though the symmetry inherent in $\CPTC$ programs is in
some sense different from the symmetry of symmetric circuits. It might be more
natural then to ask if $\CPTC$ can be characterised using families of circuits
with a weakened symmetry requirement.

% The most obvious next step would be to try and develop a (natural) circuit
% characterisation of $\CPTC$. There are a number of trivial approaches that can
% be used to define a circuit characterisation for $\CPTC$. For example, if $Q$
% is a decision problem in $\CPTC$ that is complete under first-order
% reductions, and introduce define a circuit over a basis that includes
% functions computing that decide that decide this problem. However, if a more
% nau

% It would be natural to ask if there is a class of circuits, perhaps symmetric
% circuits defined over an extension of the majority basis, that would suffice
% to give a circuit characterisation of $\CPT$ or $\CPTC$ ($\CPT$ with
% counting). We recall that $\CPTC$ is one of the two candidate logics for
% capturing polynomial-time, and as such there is significant interest in
% understanding its expressive power. At present, very little is known about
% $\CPTC$, and there are no known promising techniques for proving lower bounds.
% Since circuits families seem easier to analysis then $\CPTC$ programs, a
% circuit characterisation of $\CPTC$ would provide us with a very promising for
% studying the expressive power of $\CPTC$, perhaps allowing us to apply
% techniques for finding circuit lower bounds to the problem of deriving
% inexpressibly results for $\CPTC$. As a starting point we might ask if there
% are any interesting extensions of the majority basis by non-symmetric
% functions such that $\PT$-uniform symmetric circuits over such a basis is
% comparable to $\CPT$ -- i.e.\ known to be at least or at most as expressive as
% $\CPT$. We know that any fragment of $\CPTC$ given by placing a constant bound
% on the rank of sets used in computation is at least as expressive as $\FPC$
% (see~\cite{}), and hence at least as expressive as $\PT$-uniform symmetric
% circuits with symmetric gates. We may ask if for each fragment of $\CPTC$ with
% a bound $r$ on the rank of the sets used, there is a non-trivial class of
% circuits $\mathcal{C}_r$ (perhaps a family of symmetric circuits over some
% basis) such that the fragment of $\CPT$ with bound $r$ is at most as
% expressive as the $\PT$-uniform families of circuits from $\mathcal{C}_r$. It
% is worth noting that, since we now have a circuit characterisation for $\FPR$,
% finding a circuit characterisation for $\CPT$ or $\CPTC$ (or showing none
% exists) may help clarify the relationship between $\CPTC$ and $\FPR$.

% We have proposed a new circuit model, presented a characterisation of $\FPR$
% in terms of $\PT$-uniform families of a particular class of circuits, and
% suggested how we might generalise these results in order to develop
% circuit-based characterisations of other logics -- hopefully even $\CPT$ or
% $\CPTC$. The natural next step would be to develop techniques for proving
% lower bounds for $\PT$-uniform circuit families. A general approach for
% finding lower bounds for families of symmetric circuits over some basis may be
% very powerful, particularly if it turns out to be possible to characterise a
% broad class of extensions of $\FP$ using $\PT$-uniform families of transparent
% symmetric circuits. The usual approach for establishing inexpressibly results
% for logics involves first defining a pebble game for the logic. Dawar~\cite{}
% has shown that the pebble game of Hella~\cite{} can be directly linked with
% symmetric circuits with symmetric gates via a circuit-based argument that
% makes no reference to logic. Dawar and Holm~\cite{} have defined a pebble game
% for $\FPR$, and it would be interesting to know if a similar link can be
% established between this pebble game and transparent symmetric rank-circuits.
% It may be possible to generalise these arguments and establish for each basis
% $\BB$ a corresponding pebble game for transparent symmetric circuits defined
% over $\BB$.

% The circuit model we have developed is novel, and as such there are many basic
% questions of interest. We do not know, for example, whether for each $a \in
% \nats$ there is a symmetric circuit with symmetric gates where each gate has
% fan-in less than $a$ that computes $\NAND[a+1]$. For a basis $\BB$ let $Q_\BB$
% be the set of all queries computed by $\PT$-uniform families of transparent
% symmetric circuits defined over $\BB$. We have that $Q_{\BM} \subsetneq
% Q_{\RB}$. It follows from Theorem~\ref{} that if $\BB$ is a symmetric basis
% then $Q_\BB \subseteq Q_{\BM}$. It follows that $Q_{\BM}$ is a basis including
% structured functions defined over a vocabulary with two sorts that defines a
% circuit model strictly more powerful than any defined by basis of that only
% include functions with one sort.s

% defines a strictly more powerful class of circuits than

% It is not hard to show that $Q_\BB \subseteq Q_{\BM}$ for any matrix-symmetric
% basis $Q_\BB$ consisting of polynomial-time computable functions if, and only
% if, $\FPR$ captures $\PT$.

% We know from Theorem~\ref{} that the majority basis is `complete' for
% symmetric bases in the sense that any query computed by a $\PT$-uniform family
% of symmetric circuits defined over a symmetric basis there is a $\PT$-uniform
% family of symmetric circuits over the majority basis that computes the same
% query. It is easy to see that if

% Similarly, we do not know if for $r, a, p \in \nats$, with $p$ prime, the
% function $\RANK^r_p [a][a]$ is computable by a symmetric circuit with
% symmetric gates.

\end{document}

### 2018-04-06 sections/introduction.tex
Most recent changes to the conclusion, with edits all around the paper.

The study of extensions of fixed-point logics plays an important role
in the field of descriptive complexity theory.  In particular,
fixed-point logic with counting ($\FPC$) has become a reference logic
in the search for a logic for polynomial-time
(see~\cite{Dawar-siglog}).  In this context, Anderson and
Dawar~\cite{AndersonD17} provide an interesting characterization of
the expressive power of $\FPC$ in terms of circuit complexity.  They
show that the properties expressible in this logic are exactly those
that can be decided by polynomially-uniform families of circuits (with
threshold gates) satisfying a natural \emph{symmetry} condition.  Not
only does this illustrate the robustness of $\FPC$ as a complexity
class within $\PT$ by giving a distinct and natural characterization
of it, it also demonstrates that the techniques for proving
inexpressibility in the field of finite model theory can be understood
as lower-bound methods against a natural circuit complexity class.
This raises an obvious question (explicitly posed in the concluding
section of~\cite{AndersonD17}) of how to obtain circuit
characterizations of logics more expressive than $\FPC$, such as
choiceless polynomial time (CPT) and fixed-point logic with rank
($\FPR$).  It is this last question that we address in this paper.

@@

Another crucial ingredient in the proof of Anderson and Dawar is that
we can eliminate redundancy in the circuit $C_n$ by making it
\emph{rigid}.  That is, we can ensure that the \emph{only}
automorphisms of $C_n$ are those that are induced by permutations of
$[n]$.  Here we face the difficulty that identifying the symmetries
and eliminating redundancy in a circuit that involves gates computing
$\tau$-invariant functions requires us to solve the isomorphism
problem for $\tau$-structures.  This is a hard problem (or, at least,
one that we do not know how to solve efficiently) even when the
$\tau$-structures are $0$-$1$-matrices.  We overcome this difficulty
by placing a further restriction on circuits that we call
\emph{transparency}.  Circuits satisfying this condition have the
property that their lack of redundancy is transparent.  We explore
this condition and its consequences in Section~\ref{sec:transparent},
where we also demonstrate that something like this requrement is
necessary.  Natural weakenings of this requirement yield circuits in
which deciding symmetry seems hard.

>>>>

The study of extensions of fixed-point logics plays an important role
in the field of descriptive complexity theory.  In particular,
fixed-point logic with counting ($\FPC$) has become a reference logic
in the search for a logic for polynomial-time
(see~\cite{Dawar-siglog}).  In this context, Anderson and
Dawar~\cite{AndersonD17} provide an interesting characterisation of
the expressive power of $\FPC$ in terms of circuit complexity.  They
show that the properties expressible in this logic are exactly those
that can be decided by polynomially-uniform families of circuits (with
threshold gates) satisfying a natural \emph{symmetry} condition.  Not
only does this illustrate the robustness of $\FPC$ as a complexity
class within $\PT$ by giving a distinct and natural characterization
of it, it also demonstrates that the techniques for proving
inexpressibility in the field of finite model theory can be understood
as lower-bound methods against a natural circuit complexity class.
This raises an obvious question (explicitly posed in the concluding
section of~\cite{AndersonD17}) of how to obtain circuit
characterizations of logics more expressive than $\FPC$, such as
choiceless polynomial time (CPT) and fixed-point logic with rank
($\FPR$).  It is this last question that we address in this paper.

@@

Another crucial ingredient in the proof of Anderson and Dawar is that
we can eliminate redundancy in the circuit $C_n$ by making it
\emph{rigid}.  That is, we can ensure that the \emph{only}
automorphisms of $C_n$ are those that are induced by permutations of
$[n]$.  Here we face the difficulty that identifying the symmetries
and eliminting redundancy in a circuit that involves gates computing
$\tau$-invariant functions requires us to solve the isomorphism
problem for $\tau$-structures.  This is a hard problem (or, at least,
one that we do not know how to solve efficiently) even when the
$\tau$-structures are $0$-$1$-matrices.  We overcome this difficulty
by placing a further restriction on circuits that we call
\emph{transparency}.  Circuits satisfying this condition have the
property that their lack of redundancy is transparent.  We explore
this condition and its consequences in Section~\ref{sec:transparent},
where we also demonstrate that something like this requrement is
necessary.  Natural weakenings of this requirement yield circuits in
which deciding symmetry seems hard.

### 2018-04-06 sections/translating-formulas-to-cicuits.tex
Most recent changes to the conclusion, with edits all around the paper.

  \begin{remark}
    G: I have removed the section where I define equality between ground
    formulas. Do you find this adds ambiguity? I have added a little bit
    explaining what I mean by `the matrix at this point remains unchanged'. Does
    this help?
  \end{remark}

@@

  %We also relabel the bound variables in $\lambda(\vec{x})$ such that there
  % are
  % no two sub-formulas $\psi$ and $\psi'$ such that $\psi$ and $\psi'$ are
  % equal
  % up to a renaming of those variables that are bound in $\psi$ or $\psi'$.

@@

We now use Lemma~\cite{lem:translating-FOrk} to prove that for every $\FPR$
formula there is a corresponding $\PT$-uniform family of transparent symmetric
rank-circuits that decides the same query.

@@

  \begin{remark}
    A: Need more specific references than the above.
  \end{remark}

>>>>

  % \begin{remark}
  %   G: I have removed the section where I define equality between ground
  %   formulas. Do you find this adds ambiguity? I have added a little bit
  %   explaining what I mean by `the matrix at this point remains unchanged'.
  %   Does
  %   this help?
  % \end{remark}

@@

  % We also relabel the bound variables in $\lambda(\vec{x})$ such that there
  % are
  % no two sub-formulas $\psi$ and $\psi'$ such that $\psi$ and $\psi'$ are
  % equal
  % up to a renaming of those variables that are bound in $\psi$ or $\psi'$.

@@

We now use Lemma~\ref{lem:translating-FOrk} to prove that for every $\FPR$
formula there is a corresponding $\PT$-uniform family of transparent symmetric
rank-circuits that decides the same query.

@@

  % \begin{remark}
  %   A: Need more specific references than the above.
  % \end{remark}
